---
title: "Homework 04"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(gapminder)
```

```{r}
str(gapminder)
my_data <- gapminder


## Need year, country, life expectancy for 2 countries
my_cameroon <- my_data %>% 
  filter(country == 'Cameroon') %>% 
  select(year, lifeExp) %>% 
  rename(lifeExp_cameroon = lifeExp)
  
my_france <- my_data %>% 
  filter(country == 'France')%>% 
  select(year, lifeExp)%>% 
  rename(lifeExp_france = lifeExp)

new_table <- left_join(my_cameroon, my_france, by = "year")

knitr::kable(new_table)

p <- ggplot(new_table, aes(x = lifeExp_cameroon, y = lifeExp_france))
p + geom_point()


## do this using spread instead

my_data_spread <- gapminder %>% 
  filter(country == "Cameroon" | country == "France") %>% 
  select(year, country, lifeExp) %>% 
  spread(key = "country", value = "lifeExp")

knitr::kable(my_data_spread)

p <- ggplot(my_data_spread, aes(x = Cameroon, y = France, label = year, color = year))
p + geom_point() +
  geom_text(check_overlap = TRUE, nudge_y = 0.5)


  
```

Activity #3
Compute mean life expectancy for all combinations of continent and year. Reshape to have one row per year and one variable for each continent.
```{r}
my_data2 <- my_data %>% 
  group_by(continent, year) %>% 
  summarise(mean_life_exp = mean(lifeExp))

p2 <- ggplot(my_data2, aes(x = year, y = mean_life_exp))
p2 + geom_point(aes(color = continent)) 

p2 + geom_point(aes(color = continent)) +
  facet_wrap(~continent)

my_data2_table <- my_data2 %>% 
  spread(key = continent, value = mean_life_exp)

knitr::kable(my_data2_table)
```

Join, Merge, Lookup

Activity 2: Make your own cheatsheet

```{r}
## Make 3 data frames

imdb <- "
                          movie,      director,              studio,     budget,    genre, country
                     Fight Club, David Fincher,                 Fox,   63000000,     drama,   USA
One Flew Over the Cuckoo's Nest,  Milos Forman,       Fantasy Films,    4400000,     drama,   USA
                  The Searchers,     John Ford,         Warner Bros,    3750000, adventure,   USA
             American History X,     Tony Kaye,     New Line Cinema,    7500000,     crime,   USA
             The Maltese Falcon,   John Huston,         Warner Bros,     375000,   mystery,   USA
                Cinema Paradiso, Giuseppe Tornatore,  Cristaldifilm,    5000000,     drama, Italy
               Django Unchained, Quentin Tarantino,       Weinstein, 100000000,  western,   USA
              Princess Mononoke, Hayao Miyazaki,             DENTSU,   23151960, adventure, Japan
                   Citizen Kane,   Orson Welles, RKO Radio Pictures,     839727,   mystery,   USA
                        Amadeus,  Milos Forman,                AMLF,   18000000,   biography,   USA
          To Kill a Mockingbird, Robert Mulligan, Universal International, 2000000, crime, USA
                           Room, Lenny Abrahamson,  Element Pictures,  13000000,     drama, Ireland
                      True Grit, Coen brothers,   Paramount Pictures,   38000000,   western,   USA
                       Deadpool,    Tim Miller,                  Fox,  58000000,    comedy,   USA
             The Imitation Game, Morten Tyldum,  Black Bear Pictures,  14000000, biography, UK
                   Donnie Darko, Richard Kelly,       Pandora Cinema,   4500000, drama, USA
                The Truman Show, Peter Weir,      Paramount Pictures,  60000000, comedy, USA
                         Ip Man, Wilson Yip,  Beijing ShengShi HuaRei, 11715578, biography, HK
                         Gandhi, Richard Attenborough,  International Film Investors, 22000000, biography, UK
    The Cabinet of Dr. Caligari, Robert Wiene, Decla-Bioscop AG, 18000, fantasy, Germany
               The way way back,  Nat Faxon, Sycamore Pictures, 5000000, comedy, USA
                       Clueless, Amy Heckerling,  Paramount Pictures, 12000000, comedy, USA
"

imdb <- read_csv(imdb, trim_ws = TRUE, skip = 1)

knitr::kable(imdb)

viewed <- "
                          movie,  seen, year_released
                     Fight Club,  TRUE, 1999
One Flew Over the Cuckoo's Nest,  TRUE, 1975
                  The Searchers,  TRUE, 1956
             American History X,  TRUE, 1998
             The Maltese Falcon, FALSE, 1941
                Cinema Paradiso,  TRUE, 1988
               Django Unchained,  TRUE, 2012
              Princess Mononoke,  TRUE, 1997
                   Citizen Kane,  TRUE, 1941
                        Amadeus,  TRUE, 1984
          To Kill a Mockingbird,  TRUE, 1962
                           Room, FALSE, 2015
                      True Grit,  TRUE, 2010
                       Deadpool,  TRUE, 2016
             The Imitation Game,  TRUE, 2014
                   Donnie Darko, FALSE, 2001
                The Truman Show,  TRUE, 1998
                         Ip Man, FALSE, 2008
                         Gandhi, FALSE, 1982
    The Cabinet of Dr. Caligari,  TRUE, 1920
               The way way back,  TRUE, 2013
                       Clueless,  TRUE, 1995
"
viewed <- read_csv(viewed, trim_ws = TRUE, skip = 1)

knitr::kable(viewed)

studios <- "
        studio, year_est
Weinstein, 2005
Universal International, NA
Universal, 1912
Fox, 1935
Paramount Pictures, 1912
Element Pictures, NA
Enoki Films, 1975
Dreamworks, 1994
"
studios <- read_csv(studios, trim_ws = TRUE, skip = 1)

knitr::kable(studios)

```

inner_join(x, y): Return all rows from x where there are matching values in y, and all columns from x and y. This is a mutating join
```{r}
ij_is <- inner_join(imdb, studios)
ij_si <- inner_join(studios, imdb)

knitr::kable(ij_is)
knitr::kable(ij_si)


ij_iv <- inner_join(imdb, viewed)
knitr::kable(ij_iv)

```


semi_join(x, y): Return all rows from x where there are matching values in y, keeping just columns from x. A semi join differs from an inner join because an inner join will return one row of x for each matching row of y, where a semi join will never duplicate rows of x. This is a filtering join.
```{r}
## only returns movies made by a studio that is in the studio table
sj_is <- semi_join(imdb, studios)
knitr::kable(sj_is)

## only returns studios that also made a film in the imdb table
sj_si <- semi_join(studios, imdb)
knitr::kable(sj_si)
```


left_join(x, y): Return all rows from x, and all columns from x and y. If there are multiple matches between x and y, all combination of the matches are returned. This is a mutating join.
```{r}
## returns all rows from "movie" in imdb and all columns from imdb and viewed
lj_iv <- left_join(imdb, viewed)
knitr::kable(lj_iv)

## returns all rows from "movie" in viewed and all columns from imdb and viewed
lj_vi <- left_join(viewed, imdb)
knitr::kable(lj_vi)

## returns all rows from "studio" in imdb and all columns from imdb and studios
lj_is <- left_join(imdb, studios)
knitr::kable(lj_is)

## returns all rows from "studio" in studios (including every instance that it appears in imdb) and all columns from imdb and studios
lj_si <- left_join(studios, imdb)
knitr::kable(lj_si)

```


anti_join(x, y): Return all rows from x where there are not matching values in y, keeping just columns from x. This is a filtering join.
```{r}
## returns all rows from "studio" in imdb that do not appear in studios table. Only keeps columns from imdb
aj_is <- anti_join(imdb, studios)

## returns all rows from "studio" in studios that do not appear in imdb table. Only keeps columns from studios
aj_si <- anti_join(studios, imdb)
knitr::kable(aj_si)
```

Full join
full_join(x, y): Return all rows and all columns from both x and y. Where there are not matching values, returns NA for the one missing. This is a mutating join.
```{r}
fj_is <- full_join(imdb, studios)
knitr::kable(fj_is)

fj_iv <- full_join(imdb, viewed)
```

I made these data frames with different classes of data (logical, integer, character) so that I can play around with graphing and manipulating objects later. In hindsight, I think I should make another data frame with less overlap, and fewer rows, so that the cheatsheet is more functional. I just really like movies, so it was hard to stop adding to the list ;)

