---
title: "Homework 05"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load packages:
```{r}
library(tidyverse)
library(forcats)
library(gapminder)
```

# Factor management:
## Drop Oceania.
```{r}
## Explore gapminder factor continent
str(gapminder)
str(gapminder$continent)
levels(gapminder$continent)
nlevels(gapminder$continent)
fct_count(gapminder$continent)

## Drop Oceania using 2 methods:
## method 1:
my_levels <- c("Africa", "Americas", "Asia", "Europe")
gapminder %>% 
  filter(continent %in% my_levels)

## method 2:
my_gap_drop_oceania <- gapminder %>% 
  filter(continent != "Oceania")

## number of rows in gapminder
nrow(gapminder)
## number of rows now that oceania is gone
nrow(my_gap_drop_oceania)

## remove unused factor levels:
nlevels(my_gap_drop_oceania$continent) # still have 5 levels

my_gap_drop_oceania$continent %>% 
  fct_drop() %>% 
  levels() # removes levels from stand-alone factor

drop_oceania <- my_gap_drop_oceania %>% 
  droplevels()
nlevels(drop_oceania$continent) # now there are 4 levels
```

## Reorder the levels of country or continent.
### Change order of country based on mean life expectancy:
```{r}
## level order for country before changing the factor level based on mean life expectancy:
drop_oceania$country %>% 
  levels() %>% 
  head()

## ...and now after re-ordering:
fct_reorder(drop_oceania$country, drop_oceania$lifeExp, mean) %>% 
  levels() %>% 
  head()

## note: fct_reorder doesn't actually change anything in the data frame...that's what arrange() does
write_csv(drop_oceania, "drop_oceania.csv")
drop_oceania <- read_csv("drop_oceania.csv")

## test using arrange() vs. fct_reorder()
arr_drop_oceania <- drop_oceania %>% 
  arrange(lifeExp)

arr_drop_oceania1 <- drop_oceania %>% 
  group_by(country, continent) %>% 
  summarise(mean_life_exp = mean(lifeExp)) %>% 
  arrange(mean_life_exp)

```
## Apologies in advance for how long these graphs are. I don't think it's the best way to look at these data, but maybe it's the best way to look at arrange() vs. forcats?
```{r, fig.width = 10, fig.height = 25}
## Solely using arrange() creates a messy plot that does not mirror the order of "country" by "mean_life_exp" the new tibble created
p1 <- ggplot(arr_drop_oceania1, aes(x = mean_life_exp, y = country)) +
  geom_point(aes(color = continent))
p1

## Have to use reorder() to actually plot in the same order as arr_drop_oceania1, which was created using arrange().
## I don't know why yet.
p2 <- ggplot(arr_drop_oceania1, aes(x = mean_life_exp,
                      y = reorder(country, mean_life_exp)))
p2 + geom_point(aes(color = continent))

## Using arrange() without grouping and summarizing first.
p3 <- ggplot(arr_drop_oceania, aes(x = lifeExp, y = country))
p3 + geom_point(aes(color = continent))

## Using reorder() in conjunction with arrange() creates the same graph we can make more simply using forcats. See next graph.
p4 <- ggplot(arr_drop_oceania, aes(x = lifeExp, y = reorder(country, lifeExp)))
p4 + geom_point(aes(color = continent))


## I wanted to play with factor levels more here --- will come back to if I have time.
# arr_drop_oceania2 <- my_gap_drop_oceania %>% 
#   arrange(lifeExp)
# p6 <- ggplot(my_gap_drop_oceania, aes(x = lifeExp, y = reorder(country, lifeExp)))
# p6 + geom_point(aes(color = continent))

```

### Using forcats
```{r, fig.width = 10, fig.height = 25}
p5 <- ggplot(drop_oceania, aes(x = lifeExp, 
                               y = fct_reorder(drop_oceania$country, drop_oceania$lifeExp, mean)))
p5 + geom_point(aes(color = continent))

```

# Experimenting further with File I/O
```{r}
my_gap <- gapminder
gap_pop <- my_gap %>% 
  group_by(country, continent) %>%
  summarize(pop = max(pop)) %>% 
  ungroup()

gap_pop <- gap_pop %>% 
  mutate(country = fct_reorder(country, pop))
levels(gap_pop$country) %>% 
  head()

## I want to save the manipulations to factor level order that I made, so am saving the R object as 
## opposed to writing it as a csv or txt file
saveRDS(gap_pop, "gap_pop.rds")
rm(gap_pop) # remove the object before trying to read it back in

gap_pop <- readRDS("gap_pop.rds")
levels(gap_pop$country) %>% 
  head()
## you can see here that the factor order survives re-entry to R

## Proof of concept graph to show myself that changes by fct_reorder() translate to the figure.
## I'm leaving it squished, so that the markdown file isn't ridiculously long.
ggplot(gap_pop, aes(x = pop, y = country)) + geom_point(aes(color = continent)) + scale_x_log10()


```

```{r, fig.width=10, fig.height=25}
## More playing with factors and figures
my_gap <- gapminder
gap_pop <- my_gap %>% 
  group_by(country, continent, lifeExp, year) %>%
  summarize(pop = max(pop)) %>% 
  ungroup()

gap_pop <- gap_pop %>% 
  mutate(country = fct_reorder(country, pop))
levels(gap_pop$country) %>% 
  head()
ggplot(gap_pop, aes(x = pop, y = country)) + geom_point(aes(color = year, size = lifeExp, alpha = 1/10)) + scale_x_log10()
```

# Visualization design
```{r}
## here is a plot I made in homework 2 looking at the relationship between life expectancy and GDP per capita
ggplot(gapminder, aes(x = gdpPercap, y = lifeExp)) +
  scale_x_log10() +
  geom_point(aes(color = continent, alpha = 1/3)) +
  geom_smooth()

## let's find a better way to show these data
## Multiply pop by -1 to order from largest to smallest. Removed Oceania because not enough data points
new_gdp_data <- gapminder %>% 
  filter(continent != "Oceania") %>% 
  droplevels() %>% 
  mutate(country = reorder(country, -1*pop, median)) %>% 
  arrange(year, country)  

test_year <- 2007

new_gdp_plot <- new_gdp_data %>% 
  filter(year == test_year)

p7 <- ggplot(new_gdp_plot, aes(x = gdpPercap, y = lifeExp)) + 
  scale_x_log10(limits = c(200, 75000))

# p7 + geom_point(aes(size = pop), pch = 23)

s <- p7 + geom_point(aes(size = pop), pch = 21, show.legend = FALSE) + scale_size_continuous(range = c(1,45))

s <- s + facet_wrap(~continent) + ylim(c(38,100)) + aes(alpha = 1/3, fill = country) +
  scale_fill_manual(values = country_colors)
s
## size of the circle is relative to population size
```

# Use ggsave()
```{r}
ggsave("new_gdp_plot.png", s)

```

# Progress report
I struggled to remember what was going on behind the scenes with the factor reordering (i.e. what was actually changed and what wasn't, etc...). I think I have a tenuous hold on it right now. 

I am still not totally clear on when to use dput() and dget().

I think manipulating data visualization using ggplot seems pretty straightforward, but I wish I had more time to really explore it. I am sure that I have barely scratched the surface.