---
title: "Homework 03: Use dplyr to manipulate and explore data"
author: "Dre"
date: "October 1, 2016"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(gapminder)
library(tidyverse)
library(RColorBrewer)
```

```{r}
summary(gapminder)
## str(gapminder)
## glimpse(gapminder)

```

Get the maximum and minimum GDP per capita for each continent
```{r}
## remname gapminder data frame
my_data <- gapminder

## Get the maximum and minimum GDP per capita for each continent
## Telling R that you want to group the data by levels of a factor (in this case continent)
## Feed that into the summarize command where you create 2 new variables that provide the max or min of
## gdpPercap within a given group (continent), and then combines the results into a new tibble 
## gdp_per_cap_summary

gdp_percap_summary <- my_data %>% 
  group_by(continent) %>%
  summarize(max_gdp_percap = max(gdpPercap), min_gdp_percap = min(gdpPercap))

knitr::kable(gdp_percap_summary)

gdp_summary1 <- data.frame(gdp_percap_summary$continent, gdp_percap_summary$max_gdp_percap)
gdp_summary2 <- data.frame(gdp_percap_summary$continent, gdp_percap_summary$min_gdp_percap)
gdp_summary1 <- gdp_summary1 %>% 
  mutate(stat_done = rep(c("max"),nrow(gdp_percap_summary))) %>% 
  rename(continent = gdp_percap_summary.continent,
         gdp_percap = gdp_percap_summary.max_gdp_percap)

gdp_summary2 <- gdp_summary2 %>%
  mutate(stat_done = rep(c("min"),nrow(gdp_percap_summary))) %>% 
  rename(continent = gdp_percap_summary.continent,
         gdp_percap = gdp_percap_summary.min_gdp_percap)

gdp_percap_graph <- rbind(gdp_summary1, gdp_summary2)
gdp_percap_graph$stat_done <- as.factor(gdp_percap_graph$stat_done)

gdp_percap_graph %>% 
  as_tibble()

## rearranged tibble so it now looks like this:
knitr::kable(gdp_percap_graph)

f_colors = c("deepskyblue", "deepskyblue4")
names(f_colors) <- levels(gdp_percap_graph$stat_done)

p <- ggplot(gdp_percap_graph, aes(x = continent, y = gdp_percap, fill = stat_done)) + scale_y_log10()
p + geom_bar(stat = "identity", position=position_dodge()) + 
  scale_fill_manual(values = f_colors)


## Looking at the data differently
p <- ggplot(gdp_percap_graph, aes(x = stat_done, y = gdp_percap, group = continent, color = continent)) + scale_y_log10()
p + geom_line(size = 1) + geom_point(aes(color = continent))

p + geom_line(size = 1) + geom_point(aes(color = continent)) +
  facet_wrap(~continent)


```

Look at the spread of GDP per capita for all continents 
```{r}
gdp_by_continent <- my_data %>% 
  group_by(continent, country) %>% 
  summarise(mean_gdp = mean(gdpPercap), sd_gdp = sd(gdpPercap))

knitr::kable(gdp_by_continent)

ggplot(gdp_by_continent, aes(x = continent, y = mean_gdp, color = continent)) + scale_y_log10() +
  geom_boxplot(outlier.colour = "black") +
  geom_jitter(position = position_jitter(width = 0.3, height = 0), alpha = 1/3, size = 2)

gdp_spread_continent <- my_data %>% 
  group_by(continent) %>% 
  summarise(mean_gdp = mean(gdpPercap), sd_gdp = sd(gdpPercap), max_gdp = max(gdpPercap), min_gdp = min(gdpPercap))

knitr::kable(gdp_spread_continent)

```

How is life expectancy changing over time on different continents?
```{r}
life_exp_by_continent <- my_data %>% 
  group_by(continent, year) %>% 
  summarize(mean_life_exp = mean(lifeExp))
  
knitr::kable(life_exp_by_continent)

p <- ggplot(life_exp_by_continent, aes(x = year, y = mean_life_exp, color = continent)) 
p + geom_point() +
  geom_smooth()

p + geom_point() +
  geom_smooth() + 
  facet_wrap(~continent) 

```

I used a lot of the online material linked to through the course website and the R cheatsheets to help remind me which commands would do what I was trying to achieve. I wish I had more time to make more interesting graphs.